---
title: "CASEstudy2DS"
author: "Solange GarciadeAlford"
date: "December 1, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Science - Case Study 2

##Executive Summary

##Talent management is at the forefront of CDF’s business strategy.  One of the most valuable assets at CDF are the Employees that form the company.  

##CDF’s decision to analyze the company’s attrition is to ensure CDF continues to maintain its place as the leader in the industry.  Each one of CDF’s Employees is key to the success of the enterprise and to gain a competitive edge over its competition.

##DDSAnalytics was engaged by CDF to provide business intelligence to help retain its employees.  

##This analysis covers:
##   Attrition Trends
##   Top factors that contribute to CDF’s employee turnover
##   Attrition Prediction Model


```{r}


options(repos="https://cran.rstudio.com" )
install.packages("caret", dependencies = TRUE)
library(caret)

install.packages("class::knn")
library(mlr)

library(class)
require(class)
install.packages("fpp")
library(fpp2)
install.packages("dygraphs")
library(dygraphs)
library(xts)
library(lubridate)

library(RColorBrewer)
##display.brewer.all()
cols<-brewer.pal(n=4,name="Set1")

## IMPORT DATA
## training data set 
trainingdata <- read.csv("C:/Users/solan/CaseStudy2-data.csv")
## Validation set 
testdata <- read.csv("C:/Users/solan/CaseStudy2Validation.csv")


## Pull Attrition data into new data frame
atty <- data.frame(trainingdata[1:1170,c(1,2,3,25,26,27,29,30,31,32,33,34,35,36)])
attyfinal <- atty[!(atty$Attrition == "No"),]




## Pull of Current employee data into new data frame
attn <- data.frame(trainingdata[1:1170,c(1,2,3,25,26,27,29,30,31,32,33,34,35,36)])
attnfinal <- attn[!(attn$Attrition == "Yes"),]



## Encoding of response variable since the response variable is a categorical variable
trainingdata = within(trainingdata, {
    Attrition1 = ifelse(Attrition == 'Yes', 1, 0)   
})
testdata = within(testdata, {
    Attrition1 = ifelse(Attrition == 'No', 1, 0)
})


### Encoding of explanatory variables

## Encoding Gender
trainingdata = within(trainingdata, {
    gender1 = ifelse(Gender == 'Female', 1, 0)   
})
testdata = within(testdata, {
    gender1 = ifelse(Gender == 'Female', 1, 0)
})
## Encoding Department
trainingdata$Department1[trainingdata$Department == 'Human Resources'] <- 1
trainingdata$Department1[trainingdata$Department == 'Research & Development']<- 2
trainingdata$Department1[trainingdata$Department == 'Sales'] <- 3

testdata$Department1[testdata$Department == 'Human Resources'] <- 1
testdata$Department1[testdata$Department == 'Research & Development']<- 2
testdata$Department1[testdata$Department == 'Sales'] <- 3


## Trends - Scatter Plots

plot(attnfinal$ID, attnfinal$YearsAtCompany, main="  Current Employees - Years at the Company", 
  	xlab="Employee", ylab="Employee's Years At Company", pch=20, col="Green")


plot(attnfinal$ID, attnfinal$Age, main="  Current Employees by Age", 
  	xlab="Employee by ID ", ylab="Curren Employees by Age", pch=20, col="orange")


plot(attyfinal$TotalWorkingYears, main="  Attrition - Total Working Years", 
  	xlab="Employees", ylab="Employee's Years At Company", pch=20, col="blue")


## Fitting model on the trainingset
Model1_fit = lm(Attrition1 ~ Age + YearsAtCompany + YearsSinceLastPromotion + MonthlyIncome + Department1, data=trainingdata)
summary(Model1_fit)




Model1_res = resid(Model1_fit)

plot(trainingdata$Age, Model1_res, ylab="Residuals", xlab="Age", main="Age Residuals Plot", col="blue") 
abline(0, 0)                  # the horizontal line


## Predicting Attrition - CREATE THE TARGET OBJECTS
dfTrainData <- data.frame(trainingdata[1:1170,c(1,2,7,15,20,27,29,30,32,33,35,36,39,38)])
train_target <- as.factor(trainingdata[1:1170, 3])     ## target must be saved as a factor for knn function
length(train_target)

dfTestData <- data.frame(testdata[1:300,c(1,2,7,15,20,27,29,30,32,33,35,36,39,38)])
test_target <- as.factor(testdata[1:300, 3])           ## target must be saved as a factor for knn function
length(test_target)

## build classification Models and confusion tables
dfTrainM1 <- knn(train=dfTrainData, test=dfTestData, cl=train_target, k=11)
## Comparisons models only
dfTrainM2 <- knn(train=dfTrainData, test=dfTestData, cl=train_target, k=5)
dfTrainM3 <- knn(train=dfTrainData, test=dfTestData, cl=train_target, k=8)
dfTrainM4 <- knn(train=dfTrainData, test=dfTestData, cl=train_target, k=9)
dfTrainM5 <- knn(train=dfTrainData, test=dfTestData, cl=train_target, k=10)
## Confusion Tables
table(test_target, dfTrainM1)
confusionMatrix(dfTrainM1, test_target)
table(test_target, dfTrainM2)
confusionMatrix(dfTrainM2, test_target)
## comparison purpose only
table(test_target, dfTrainM2)
table(test_target, dfTrainM3)
table(test_target, dfTrainM4)
table(test_target, dfTrainM5)


## creation of dfTrain and dfVal data frame
###IMPORTANT NOTE ----I was not able to get rid of the NA so I had to use a reduced file####
dfTrain <- data.frame(trainingdata[1:1170,c(1,2,7,15,20,27,29,30,32,33,35,36,39,38)])
train_target <- as.factor(trainingdata[1:1170, 3])     ## target must be saved as a factor for knn function
length(train_target)

dfVal <- data.frame(testdata[1:300,c(1,2,7,15,20,27,29,30,32,33,35,36,39,38)])
test_target <- as.factor(testdata[1:300, 3])           ## target must be saved as a factor for knn function
length(test_target)

dfPreds1 <- knn(train=dfTrain, test=dfVal, cl=train_target, k=11)
dfPreds1
confusionMatrix(dfPreds1, test_target)

dfPreds <- data.frame(testdata$ID,dfPreds1)

write.csv(dfPreds, file="Case2PredictionsGarciadeAlford.csv")




```




